Casual Bandits

### 2016-01-19 causal-bandit/icml-2015/introduction.tex
Minor tweaks to setup

>>>>

Our framework also bears a superficial similarity to contextual bandit problems \cite{Agarwal2014} since the extra observations on non-intervened variables might be viewed as context for selecting an intervention. However, a crucial difference is that in our model the extra observations are only revealed \emph{after} selecting an intevention and cannot be used as context.

### 2016-01-19 causal-bandit/icml-2015/setup.tex
Minor tweaks to setup

A learner for a casual bandit problem is given the casual model's graph $\mathcal{G}$, a set of \emph{observable} variables $\mathcal{O} \subseteq \mathcal{X}$, a set of \emph{allowed interventions} $\mathcal{A}$, a \emph{reward variable} $Y$. 
The following game is played over $T$ rounds. 
In round $t$, the learner selects an allowed intervention $do(\vec{X}_t = \vec{x}_t)$ from $\mathcal{A}$.
Values for all non-intervened variables $\vec{X}^c_t$ are sampled from $\P{\vec{X}^c_t | do(\vec{X}_t = \vec{x}_t)}$. Finally, the learner is shown the sampled values of all the observable variables in $\vec{X}^c_t \cap \mathcal{O}$ and receives the reward $Y_t$.

@@

The causal model for this class of problems has $N$ binary variables $\{ X_1, \ldots, X_N \}$ where each $X_i \in \{0,1\}$ are independent causes of a reward variable $Y \in \R$, as shown in Figure~\ref{fig:causalStructure}.
All variables are observable (\ie, $\mathcal{O} = \{Y, X_1, \ldots, X_N\}$) and the set of allowable interventions are all size 1 interventions on a given $X_i$ variable (\ie, $\mathcal{A} = \{ do(X_i = j) \colon i \in \{1, \ldots, N\}, j \in \{0,1\}\}$).
In the farming example from the introduction, $X_1$ represent temperature with $X_1=0$ denoting low and $X_1=1$ high temperature. 
In this case, the interventions $do(X_1 = 0)$ and $do(X_1 = 1)$ might represent the use of shades or heat lamps to keep the temperature low or high, respectively.

@@

In each round of the game, the learner selects a single variable to intervene on and the value to set it to while the remainder of the variables are simultaneously set by independently biased coin flips. 
The value of all variables are then used to determine the distribution of rewards for that round.
Formally, when not intervened upon we assume that each $X_i \sim \bernoulli(q_i)$ where $\vec{q} = (q_1, \ldots, q_N) \in [0,1]^N$ so that $q_i = \P{X_i = 1}$.
The value of the reward variable is distributed as $Y = r(\vec{X}) + \eta$ where $r : \{0,1\}^N \to \R$ is an arbitrary, fixed, and unknown function and $\eta$ is $1$-subgaussian noise with a distribution that may depend on $\vec{X}$.
In the farming example, this choice of $Y$ models the fixed, noisy but unknown dependence of crop yield on the temperature, soil, and moisture.

@@

From a causal perspective the key insight is that the assumed causal structure implies that
the law of $Y_t$ given we intervene to set $X_{t,i} = j$ is the same as the law of $Y_t$ conditioned
on $X_{t,i} = j$. Formally,
\eqn {
\label{eq:observe}
\P{Y_t \leq y|do(X_{t,i} = j)} = \P{Y_t \leq y|do(),X_{t,i} = j}\,.
}
This follows from application of the do-calculus \cite{Pearl2000} to our specific causal graph. It is not the case in general. 
For example if there was a variable $X'$ that caused both $X_{t,i}$ and $Y$ (or $X_{t,i}$ and any other variable $X_{t,l}$), 
that would introduce a backdoor path from $X_{t,i} \rightarrow Y_t$ and we would have to condition on $X'$ to derive the 
interventional distribution of $Y_t$ from the observational one.

>>>>

A learner for a casual bandit problem is given the casual model's graph $\mathcal{G}$, a set of \emph{observable} variables $\mathcal{O} \subseteq \mathcal{X}$, a set of \emph{allowed interventions} $\mathcal{A}$, a \emph{reward variable} $Y$. 
The following game is then played over $T$ rounds. 
In round $t$, the learner selects an allowed intervention $do(\vec{X}_t = \vec{x}_t)$ from $\mathcal{A}$.
Values for all non-intervened variables $\vec{X}^c_t$ are sampled from $\P{\vec{X}^c_t | do(\vec{X}_t = \vec{x}_t)}$. Finally, the learner is shown the sampled values of all the observable variables in $\vec{X}^c_t \cap \mathcal{O}$ and receives the reward $Y_t$.

@@

The causal model for this class of problems has $N$ binary variables $\{ X_1, \ldots, X_N \}$ where each $X_i \in \{0,1\}$ are independent causes of a reward variable $Y \in \R$, as shown in Figure~\ref{fig:causalStructure}.
All variables are observable (\ie, $\mathcal{O} = \{Y, X_1, \ldots, X_N\}$) and the set of allowable actions are all size 0 and size 1 interventions: 
\[
	\mathcal{A} = \{do()\} \cup \{ do(X_i = j) \colon i \in \{1, \ldots, N\}, j \in \{0,1\}\}.
\]
In the farming example from the introduction, $X_1$ might represent temperature (\eg, $X_1=0$ for low and $X_1=1$ for high). 
In this case, the interventions $do(X_1 = 0)$ and $do(X_1 = 1)$ might represent the use of shades or heat lamps to keep the temperature low or high, respectively.

@@

In each round of the game, the learner either purely observes by selecting $do()$ or chooses a single variable to intervene on and the value to set it to. 
The remainder of the variables are simultaneously set by independently biased coin flips. 
The value of all variables are then used to determine the distribution of rewards for that round.
Formally, when not intervened upon we assume that each $X_i \sim \bernoulli(q_i)$ where $\vec{q} = (q_1, \ldots, q_N) \in [0,1]^N$ so that $q_i = \P{X_i = 1}$.
The value of the reward variable is distributed as $Y = r(\vec{X}) + \eta$ where $r : \{0,1\}^N \to \R$ is an arbitrary, fixed, and unknown function and $\eta$ is $1$-subgaussian noise with a distribution that may depend on $\vec{X}$.
In the farming example, this choice of $Y$ models the fixed, noisy but unknown dependence of crop yield on the temperature, soil, and moisture.

@@

From a causal perspective the key insight is that the assumed causal structure implies that
the law of $Y_t$ given we intervene to set $X_{t,i} = j$ is the same as the law of $Y_t$ conditioned
on $X_{t,i} = j$. Formally,
\eqn {
\label{eq:observe}
\P{Y_t \leq y|do(X_{t,i} = j)} = \P{Y_t \leq y|do(),X_{t,i} = j}\,.
}
This follows from application of the do-calculus \cite{Pearl2000} to the causal graph for the parallel bandits problem (Fig.~\ref{fig:causalStructure}). It is not the case in general. 
For example if there was a variable $X'$ that caused both $X_{t,i}$ and $Y$ (or $X_{t,i}$ and any other variable $X_{t,l}$), 
that would introduce a backdoor path from $X_{t,i} \rightarrow Y_t$ and we would have to condition on $X'$ to derive the 
interventional distribution of $Y_t$ from the observational one.
